### 2023-04-17 outputs/paper/paper.qmd
cleaned dates and added references

---
title: "The Shades of Racism in the Beauty Industry"
subtitle: "An exploration of inclusivity and tokenism within fashion and makeup"
author: 
  - Huda Sahaf
thanks: "Code and data are available at: https://github.com/hsahaf/Racism_Beauty_Industry.git."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper aims to investigate the history of racism in the beauty industry and how this impacts the inclusivity in the market. The complexions of models are assigned a lightness value and analyzed across 8 years. These results are compared to the foundation shade ranges across all major makeup brands across the U.S revealing a continuation of compounding racism within the industry as a whole."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
library(readr)
library(RColorBrewer)
library(scales)
```

```{r}
fashion_tones <-
  read.csv(
    file = "https://raw.githubusercontent.com/the-pudding/data/master/vogue/faces.csv", 
    skip = 0 
  )

fashion_frequency <-
  read.csv(
    file = "https://raw.githubusercontent.com/the-pudding/data/master/vogue/models.csv", 
    skip = 0 
  )

makeup_shades <-
  read.csv(
    file = "https://raw.githubusercontent.com/the-pudding/data/master/makeup-shades/shades.csv", 
    skip = 0 
  )

#### Save data ####

write_csv(fashion_tones, here::here("inputs/data/fashion_tones.csv"))

write_csv(fashion_frequency, here::here("inputs/data/fashion_frequency.csv"))

write_csv(makeup_shades, here::here("inputs/data/makeup_shades.csv"))

```

```{r}

### Clean Fashion Tones Dataset: 
cleaned_fashion_tones <-
  clean_names(fashion_tones)

cleaned_fashion_tones <-
  cleaned_fashion_tones |>
  rename(
    skintone = tone,
    lightness_value = l
  )
head(cleaned_fashion_tones)
```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

### Simplifying Dates: Fashion Tones Dataset:
cleaned_fashion_tones<-
  cleaned_fashion_tones |>
  separate(
    col = date,
    into = c("Month", "Day", "Year"),
    sep = "/"
  ) |>
  mutate(cleaned_fashion_tones, clean_date = paste(Year, Month, Day, sep = "-"))
  str(cleaned_fashion_tones$clean_date)

```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

### Clean Fashion Frequency Data Set: 
cleaned_fashion_frequency<-
  clean_names(fashion_frequency)

cleaned_fashion_frequency <-
  cleaned_fashion_frequency |>
  rename(
    skintone = tone,
    lightness_value = l,
    num_covers = n_covers
  )

```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

### Clean Makeup Shades Data Set:
cleaned_makeup_shades <-
  clean_names(makeup_shades)

cleaned_makeup_shades<-
  cleaned_makeup_shades |>
  rename(
    hue = h,
    lightness = l,
    saturation = s,
    value = v,
    hex_code = hex
  )|>
  mutate(lightness_value = lightness/100)

cleaned_makeup_shades<-
  cleaned_makeup_shades |>
  filter(group == 0|group == 1|group == 2|group == 3|group == 4)

```

```{r}
#| label: fig-fullspread
#| fig-cap: Complexion Values of All Models on Vogue Covers from 2000-2018
### Modeling Full Spread of Complexion Values on Vogue Covers from 2000-2018
cleaned_fashion_tones$clean_date <- ymd(cleaned_fashion_tones$clean_date)

entire_spread <- cleaned_fashion_tones |>
  ggplot(aes(x = lightness_value, y = clean_date)) +
  geom_point(aes(color = lightness_value < 0.4)) + 
  labs(
    x = "Lightness Value of Vogue Models",
    y = "Date of Cover",
    title = "Complexions of Models on Covers of Vogue: 2000-2018"
  ) +
  theme_classic()

datebreaks <- seq(as.Date("2001-01-01"), as.Date("2018-09-01"), by = "1 year")

print(entire_spread)
```

@fig-fullspread is a graph of the lightness value of every model that has been on the cover of Vogue from 2000 to 2018. Lightness value is

```{r}
#| message: false
#| warning: false
#| echo: false
#| tbl-cap: "Models With More than 5 Appearances on Vogue"
#| label: tbl-appearances
### Table 1: All Models With More than 5 Appearances on Vogue 
cleaned_fashion_frequency |>
  select(model, num_covers, lightness_value) |>
  group_by(num_covers) |>
  filter(num_covers > 4) |>
  kable(
    col.names = c(
      "Model",
      "Number of Covers", 
      "Lightness Value"
    ),
    digits = 2
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-whitewashing
#| fig-cap: Rihanna's Appearances on Vogue
### Graph 2: Case Study: Rihanna's 5 Appearances and the respective Complexion Values:
cleaned_fashion_tones|>
  group_by(model)|>
  filter(model=="Rihanna")|>
  ggplot(mapping = aes(x = Year, y = lightness_value, color = model)) +
  geom_point()+
  labs(
    title = "Rihanna: A Case Study in White-Washing",
    y = "Complexion Value",
    color = "Model"
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-deepest
#| fig-cap: Top 4 Deepest Complexion Models on Vogue
### Graph 3: The Top 4 Deepest Complexion Models on Vogue
cleaned_fashion_tones|>
  filter(model== "Lupita Nyongo" | model=="Rihanna"| model == "Michelle Obama" | model == "Serena Williams") |>
  ggplot(aes(y = lightness_value, color = model)) +
  geom_boxplot() +
  theme_light() +
  labs(
    y = "Lightness Value of Skintone",
    title = "Appearances of Deepest Complexion Models on Vogue"
  )
```

```{r}
#| warning: false
#| echo: false
#| label: fig-all_white
#| fig-cap: Models with Highest Number of Appearances
### Graph 4: Models with the Highest Number of Appearances on Vogue and their respective Complexion Values
cleaned_fashion_tones|>
  filter(model== "Nicole Kidman" | model=="Gisele Bundchen"| model == "Charlize Theron" | model == "Sarah Jessica Parker") |>
  ggplot(aes(y = lightness_value, color = model)) +
  geom_boxplot() +
  theme_light() +
  labs(
    y = "Lightness Value of Skintone",
    title = "Complexions of Models with Highest Number of Vogue Covers" 
  )

```

```{r}
#| message: false
#| warning: false
#| echo: false
#| tbl-cap: "US Makeup Brands: Foundation Shade Range"
#| label: tbl-shades

### Table 2: All US Brands and Their Number of Foundation Shades
cleaned_makeup_shades|>
  select(brand, product, lightness_value)|>
  group_by(brand) |>
  summarise(total_shades = n()) |>
  kable(
    col.names = c(
      "Makeup Brand",
      "Number of Foundation Shades"
      )
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-bestsellers
#| fig-cap: The Shade Distribution of Best-Selling Foundations in the U.S
### Graph 5: US Bestseller Foundations and Their Inclusivity 
cleaned_makeup_shades|>
  filter(group == 2)|>
  group_by(brand)|>
  ggplot(aes(x = lightness_value, color = brand)) +
  stat_ecdf(geom = "point") +
  theme_classic()

```

```{r}
#| warning: false
#| echo: false
#| label: fig-bipoc
#| fig-cap: The Shade Distribution of BiPOC-owned Makeup Brands
### Graph 5: BiPOC Reccs of Brands with BiPOC Founder Bestseller Foundations and Their Inclusivity 
cleaned_makeup_shades|>
  filter(group == 3)|>
  group_by(brand)|>
  ggplot(aes(x = lightness_value, color = brand)) +
  stat_ecdf(geom = "point") +
  theme_classic()
```

```{r}
#| warning: false
#| echo: false
#| label: fig-bigboys
#| fig-cap: The Shade Distribution of Largest Shade Ranges Offered in the U.S
### Graph Range of Shades With Brands with Largest Shade Ranges
cleaned_makeup_shades|>
  filter(brand == "Beauty Bakerie"|brand == "Bobbi Brown" |brand == "bareMinerals" |brand == "Bobbi Brown" | brand == "MAC"| brand == "Maybelline"|brand == "Estée Lauder"|brand == "Fenty"|brand == "Lancôme"|brand == "Make Up For Ever")|> 
  ggplot(cleaned_makeup_shades, mapping = aes(x = brand, y = lightness_value)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.3, width = 0.15, height = 0) +
  theme_minimal()

```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}

## Data Management
In order to conduct an analysis on the various parts of the beauty industry, three separate datasets were used. The Pudding is a digital publication that makes its data sets open to the public @pudd.
This paper utilizes the 
Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

Talk way more about it.

# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References