### 2023-04-19 outputs/paper/paper.qmd
Updated scripts 01 and 02

---
title: "The Shades of Racism in the Beauty Industry"
subtitle: "An exploration of inclusivity and tokenism within fashion and makeup industries"
author: 
  - Huda Sahaf
thanks: "Code and data are available at: https://github.com/hsahaf/Racism_Beauty_Industry.git."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper aims to investigate the history of racism in the beauty industry and how this impacts the inclusivity in the market. The complexions of models are assigned a lightness value and analyzed across 8 years. These results are compared to the foundation shade ranges across all major makeup brands across the U.S revealing a continuation of compounding racism within the industry as a whole."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
library(readr)
library(RColorBrewer)
library(scales)
```

```{r}
#| include: false
#| warning: false
#| message: false
# fashion_tones <-
#   read.csv(
#     file = "https://raw.githubusercontent.com/the-pudding/data/master/vogue/faces.csv", 
#     skip = 0 
#   )
# 
# fashion_frequency <-
#   read.csv(
#     file = "https://raw.githubusercontent.com/the-pudding/data/master/vogue/models.csv", 
#     skip = 0 
#   )
# 
# makeup_shades <-
#   read.csv(
#     file = "https://raw.githubusercontent.com/the-pudding/data/master/makeup-shades/shades.csv", 
#     skip = 0 
#   )
# 
# #### Save data ####
# 
# write_csv(fashion_tones, here::here("inputs/data/fashion_tones.csv")) 
# 
# write_csv(fashion_frequency, here::here("inputs/data/fashion_frequency.csv")) 
# 
# write_csv(makeup_shades, here::here("inputs/data/makeup_shades.csv"))

```

```{r}
#| include: false
#| warning: false
#| message: false
### Clean Fashion Tones Dataset: 
# cleaned_fashion_tones <-
#   clean_names(fashion_tones)
# 
# cleaned_fashion_tones <-
#   cleaned_fashion_tones |>
#   rename(
#     skintone = tone,
#     lightness_value = l
#   )
# head(cleaned_fashion_tones)
```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

### Simplifying Dates: Fashion Tones Data Set:
# cleaned_fashion_tones<-
#   cleaned_fashion_tones |>
#   separate(
#     col = date,
#     into = c("Month", "Day", "Year"),
#     sep = "/"
#   ) |>
#   mutate(cleaned_fashion_tones, clean_date = paste(Year, Month, Day, sep = "-"))
#   str(cleaned_fashion_tones$clean_date)

```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

# ### Clean Fashion Frequency Data Set: 
# cleaned_fashion_frequency<-
#   clean_names(fashion_frequency)
# 
# cleaned_fashion_frequency <-
#   cleaned_fashion_frequency |>
#   rename(
#     skintone = tone,
#     lightness_value = l,
#     num_covers = n_covers
#   )

```

```{r}
#| include: false
#| warning: false
#| echo: false
#| message: false

### Clean Makeup Shades Data Set:
# cleaned_makeup_shades <-
#   clean_names(makeup_shades)
# 
# cleaned_makeup_shades<-
#   cleaned_makeup_shades |>
#   rename(
#     hue = h,
#     lightness = l,
#     saturation = s,
#     value = v,
#     hex_code = hex
#   )|>
#   mutate(lightness_value = lightness/100)
# 
# cleaned_makeup_shades<-
#   cleaned_makeup_shades |>
#   filter(group == 0|group == 1|group == 2|group == 3|group == 4)

```

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-fullspread
#| fig-cap: Complexion Values of All Models on Vogue Covers from 2000-2018
### Modeling Full Spread of Complexion Values on Vogue Covers from 2000-2018
cleaned_fashion_tones$clean_date <- ymd(cleaned_fashion_tones$clean_date)

entire_spread <- cleaned_fashion_tones |>
  ggplot(aes(x = lightness_value, y = clean_date)) +
  geom_point(aes(color = lightness_value < 0.4)) + 
  labs(
    x = "Lightness Value of Vogue Models",
    y = "Date of Cover",
    title = "Complexions of Models on Covers of Vogue: 2000-2018"
  ) +
  theme_classic()

datebreaks <- seq(as.Date("2001-01-01"), as.Date("2018-09-01"), by = "1 year")

print(entire_spread)
```

```{r}
#| message: false
#| warning: false
#| echo: false
#| tbl-cap: "Models With More than 5 Appearances on Vogue"
#| label: tbl-appearances
### Table 1: All Models With More than 5 Appearances on Vogue 
cleaned_fashion_frequency |>
  select(model, num_covers, lightness_value) |>
  group_by(num_covers) |>
  filter(num_covers > 4) |>
  kable(
    col.names = c(
      "Model",
      "Number of Covers", 
      "Lightness Value"
    ),
    digits = 2
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-whitewashing
#| fig-cap: Rihanna's Appearances on Vogue
### Graph 2: Case Study: Rihanna's 5 Appearances and the respective Complexion Values:
cleaned_fashion_tones|>
  group_by(model)|>
  filter(model=="Rihanna")|>
  ggplot(mapping = aes(x = Year, y = lightness_value, color = model)) +
  geom_point()+
  labs(
    title = "Rihanna: A Case Study in White-Washing",
    y = "Complexion Value",
    color = "Model"
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-deepest
#| fig-cap: Top 4 Deepest Complexion Models on Vogue
### Graph 3: The Top 4 Deepest Complexion Models on Vogue
cleaned_fashion_tones|>
  filter(model== "Lupita Nyongo" | model=="Rihanna"| model == "Michelle Obama" | model == "Serena Williams") |>
  ggplot(aes(y = lightness_value, color = model)) +
  geom_boxplot() +
  theme_light() +
  labs(
    y = "Lightness Value of Skintone",
    title = "Appearances of Deepest Complexion Models on Vogue"
  )
```

```{r}
#| warning: false
#| echo: false
#| label: fig-all_white
#| fig-cap: Models with Highest Number of Appearances
### Graph 4: Models with the Highest Number of Appearances on Vogue and their respective Complexion Values
cleaned_fashion_tones|>
  filter(model== "Nicole Kidman" | model=="Gisele Bundchen"| model == "Charlize Theron" | model == "Sarah Jessica Parker") |>
  ggplot(aes(y = lightness_value, color = model)) +
  geom_boxplot() +
  theme_light() +
  labs(
    y = "Lightness Value of Skintone",
    title = "Complexions of Models with Highest Number of Vogue Covers" 
  )

```

```{r}
#| message: false
#| warning: false
#| echo: false
#| tbl-cap: "US Makeup Brands: Foundation Shade Range"
#| label: tbl-shades

### Table 2: All US Brands and Their Number of Foundation Shades
cleaned_makeup_shades|>
  select(brand, product, lightness_value)|>
  group_by(brand) |>
  summarise(total_shades = n()) |>
  kable(
    col.names = c(
      "Makeup Brand",
      "Number of Foundation Shades"
      )
  )

```

```{r}
#| warning: false
#| echo: false
#| label: fig-bestsellers
#| fig-cap: The Shade Distribution of Best-Selling Foundations in the U.S
### Graph 5: US Bestseller Foundations and Their Inclusivity 
cleaned_makeup_shades|>
  filter(group == 2)|>
  group_by(brand)|>
  ggplot(aes(x = lightness_value, color = brand)) +
  stat_ecdf(geom = "point") +
  theme_classic()

```
```{r}
#| warning: false
#| echo: false
#| label: fig-bipoc
#| fig-cap: The Shade Distribution of BiPOC-owned Makeup Brands
### Graph 5: BiPOC Recomendations of Brands with BiPOC Founder Bestseller Foundations and Their Inclusivity 
cleaned_makeup_shades|>
  filter(group == 3)|>
  group_by(brand)|>
  ggplot(aes(x = lightness_value, color = brand)) +
  stat_ecdf(geom = "point") +
  theme_classic()
```

```{r}
#| warning: false
#| echo: false
#| label: fig-bigboys
#| fig-cap: The Shade Distribution of Largest Shade Ranges Offered in the U.S
### Graph Range of Shades With Brands with Largest Shade Ranges
cleaned_makeup_shades|>
  filter(brand == "Beauty Bakerie"|brand == "Bobbi Brown" |brand == "bareMinerals" |brand == "Bobbi Brown" | brand == "MAC"| brand == "Maybelline"|brand == "Estée Lauder"|brand == "Fenty"|brand == "Lancôme"|brand == "Make Up For Ever")|> 
  ggplot(cleaned_makeup_shades, mapping = aes(x = brand, y = lightness_value)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.3, width = 0.15, height = 0) +
  theme_minimal()

```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}
## Data Management
All analysis on the data sets were done using R [@citeR]. The data was read in, cleaned, and manipulated using `dplyr`[@dan], `tidyverse` [@tidy], and `janitor` [@janice]. The graphs and tables in this paper were coded using the following packages: `ggplot2` [@gabe], `kableExtra`[@kev], `knitr` [@nat], `readr` [@ron], and `RColorBrewer`[@ryan].

## Source and Sampling
In order to conduct an analysis on the various parts of the beauty industry, three separate data sets were used. The Pudding is a digital publication that makes its data sets open to the public @pudd. A closer look is taken at Vogue using the first two data sets used for the article *Colorism in High Fashion* [@high_f]. All covers of Vogue from the year 2000 to 2018 are analyzed and categorized in these data sets. The female models on each cover are identified, and then their skin tones are filtered out of the cover, and an average color value is assigned based on the all skin tone pixels. These color values are then drained of any hue or saturation and therefore assigned a lightness value based on how light or deep the complexion of the model is, which allows for the comparison of all models on the covers of Vogue over the span of 8 years. The first data set contains the name of the model, date of their appearance on Vogue, the hex code of their skin tone, and their lightness value once the hue and saturation has been removed. This includes multiple appearances of the same model. The second data set contains the names of the models, the hex code of their skin tone, the number of covers they have made appearances in, and their average lightness value across these appearances.

The data set behind The Pudding article, *Beauty Brawl* is the third data set used in this paper [@bbshades]. It contains data about the foundation shade ranges through lightness values of multiple beauty brands around the world. For the sake of this paper, we are focusing on beauty retailers based in the US, especially since Vogue is also based in the United States.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
```

# Results
@fig-fullspread is a graph of the lightness value of every model that has been on the cover of Vogue from 2000 to 2018. Lightness value is measured on a scale from 0 to 1, 1 being extremely pale in skin tone, and 0 being extremely deep. There is a general trend present almost immediately, in that across the entire 8 year period, Vogue has consistently had lighter skinned individuals with a lightness value of 0.5 to 0.8 on their covers. It has only been after around 2009 that we see a model on Vogue with a complexion deeper than 0.4.

@tbl-appearances is a table of all the models on Vogue who have had more than 4 appearances on the covers of Vogue between 2000 and 2018. In all of the names on this list, Rihanna is the only black woman, and her average lightness value across these five appearances has been fairly midtoned, at a lightness value of 0.62.

@fig-whitewashing takes a closer look at Vogue's presentation of Rihanna's skintone on her five appearances. Vogue has been publicly accused of whitewashing Rihanna's complexion on some of her covers, and it is apparent that only after 2016 does her lightness value drop below 0.4 [@whitewashed]. While lighting and editing can account for some fluctuations in the lightness value, but it does not explain the extent of the contrast between her first three appearances and her most recent two.

@fig-deepest takes a closer look at the fluctuations in lightness values in the appearances on the models with the deepest complexions on Vogue. Rihanna has had the highest number of appearances at 5 total covers from 2000 to 2018, followed by Lupita Nyongo, at 4. Amongst these models, Lupita Nyongo has the deepest complexion and her appearances on Vogue are the most recent, her last two being in 2015 and in 2018. The fluctuation is apparent in her appearances as well, in that it is only in her most recent cover that her lightness value is shown to be deeper than 0.3.

@fig-all_white shows the fluctuation in the lightness values of the top four models with the greatest number of covers on Vogue. All four of these models are white, and all of them have made appearances in 6 or more covers. Vogue's preferences for tanned white women is clear in that in the span of 8 years, the lightness value rarely drops below 0.6.

@tbl-shades is a table of all of the major beauty brands that are based in the United States, and the number of shades in their main foundation line.

@fig-bestsellers takes a closer look at the inclusivity of the bestselling foundation lines across some of the largest beauty brands across the United States of America. It is clear that there are a greater number of shades available for people with lighter skin tones, with a lightness value of 0.6 or lighter, across all of the major brands. However, for people with a complexion that is a hair deeper than a medium skin tone, the number of shades become sparse, and the distribution spreads out as the lightness value becomes deeper than 0.5.

@fig-bipoc looks at the shade distributions of the retailers with owners that are black, indigenous, or other people of color. The distribution is much more even, and all of these beauty brands provide options for people that are on both sides of the spectrum, complexions with lightness values deeper than 0.3, as well as complexions with lightness values paler than 0.8.

@fig-bigboys looks at the distribution across the beauty brands based in the United States with a foundation line with more than 25 shades. The contrast between the number of shades provided for people with light to medium complexions (0.6-1) and the lack thereof in terms of the number of shades for people with deeper complexions (0-0.35) is apparent.

# Discussion

## First discussion point {#sec-first-point}

The same trend exists across the entire beauty industry. The deeper your complexion, the less options there are available for you. There are far less doors open for people of color in the beauty space, which causes several issues to compound. There is intersectionality at present in the communities of people impacted by this cycle of racism. Not only are these individuals at a disadvantage in society because they are women, but the issues are compounded further if they are individuals of color. These industries feed into each other, the less representation there is for black and brown women in fashion, the less options will be made available in the beauty industry and vice versa. The lack of representation creates a scarcity mindset. These women feel like they are the exception rather than the norm, and without the options available, without the same advantage of seeing people like you succeed, you are less likely to feel safe in these spaces and less confident when it comes to seeking to be in them. But the presence of black and brown women in these spaces has a significant impact.

An example of this is in the beauty industry. The 'Pull Up or Shut Up' movement in 2020 forced beauty brands to reveal how many BiPOC women were staff members and in senior leadership positions[@pullup]. It pushed for transparency in their marketing claims and resulted in a significant increase in consumer awareness, customers were now able to support brands that supported them. However, the same cannot be said for the fashion industry. The beauty industry is a commercial space that is not as difficult to break into through the rise of beauty influences, but high fashion is still considered to be a very elitist institution. BiPOC women have the power to revolutionize the beauty industry as a whole, but the lack of representation makes the path significantly harder. The compounding racism bleeds into other industries as well, such as television and film, and which hurts young girls the most. Looking into a world that has not made any room for you can be a very difficult and disheartening thing to overcome for children. Despite claims made by major institutions, this data reveals that it is only recently that black and brown women have been able to break into the industry, and it still has a long way to go in terms of actually uplifting and empowering them.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References